---
title: "Reproducible Research: Peer Assessment 1"
author: "Oscar Peña del Rio"
email: "oscar.pdr@gmail.com"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Uses the unzip() function to extract the CSV file and loads it into **activity_data** dataframe.

```{r}
unzipped_file <- unzip("activity.zip")
activity_data <- read.csv(unzipped_file)
```

## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day, omitting *NA* values. Generate histogram with the values.

```{r first_histogram, fig.width=6, fig.height=4}
steps_by_date <- aggregate(steps ~ date, activity_data, sum, na.action = na.omit)
hist(steps_by_date$steps,
     main="Histogram of steps taken per day",
     xlab="Steps interval",
     col="darkolivegreen1"
)
```

2. Calculate and report the mean and median of the total number of steps taken per day

```{r}
mean(steps_by_date$steps)
median(steps_by_date$steps)
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
